\pkg{cv}: An \proglang{R} Package for Cross-Validation of Regression Models

### 2024-01-12 scratch/notes_john/JSS-article/JSS-article-reduced-2.tex
add JSS-article-reduced-2

{\centering \includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/mpg-horsepower-scatterplot-polynomials-1}

@@

{\centering \includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/mpg-horsepower-MSE-se-1}

@@

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 735559
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.44017
bias-adjusted cross-validation criterion = 19.41622
95% CI for bias-adjusted CV criterion = (15.87764, 22.95479)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> m.auto.glm <- glm(mpg ~ poly(horsepower, 2), data=Auto)
R> boot::cv.glm(Auto, m.auto.glm)$delta
\end{CodeInput}
\begin{CodeOutput}
[1] 19.24821 19.24787
\end{CodeOutput}
\begin{CodeInput}
R> microbenchmark::microbenchmark(
+   hatvalues = cv(m.auto, k="loo"),
+   Woodbury = cv(m.auto, k="loo", method="Woodbury"),
+   naive = cv(m.auto, k="loo", method="naive"),
+   cv.glm = boot::cv.glm(Auto, m.auto.glm),
+   times=10
+ )
\end{CodeInput}
\begin{CodeOutput}
Warning in microbenchmark::microbenchmark(hatvalues = cv(m.auto, k = "loo"), :
less accurate nanosecond times to avoid potential integer overflows
\end{CodeOutput}
\begin{CodeOutput}
Unit: microseconds
      expr        min         lq       mean     median         uq        max
 hatvalues    984.287   1153.412   1160.394   1189.902   1199.414   1285.104
  Woodbury  10145.122  10213.592  10525.532  10463.385  10657.581  11476.351
     naive 216360.403 217763.218 223882.308 218184.226 219846.551 273572.951
    cv.glm 380361.674 382182.689 400581.980 386284.739 436401.540 439866.368
 neval cld
    10 a  
    10 a  
    10  b 
    10   c
\end{CodeOutput}
\end{CodeChunk}

@@

{\centering \subfloat[\label{fig:polynomial-regression-CV-graph-2-1}]{\includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/polynomial-regression-CV-graph-2-1} }\subfloat[\label{fig:polynomial-regression-CV-graph-2-2}]{\includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/polynomial-regression-CV-graph-2-2} }

@@

{\centering \subfloat[\label{fig:plot1-1}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/plot1-1} }\subfloat[\label{fig:plot1-2}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/plot1-2} }\newline\subfloat[\label{fig:plot1-3}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/plot1-3} }\subfloat[\label{fig:plot1-4}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/plot1-4} }\newline\subfloat[\label{fig:plot1-5}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/plot1-5} }

@@

{\centering \subfloat[\label{fig:cross-validation-clusters-1}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/cross-validation-clusters-1} }\subfloat[\label{fig:cross-validation-clusters-2}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced_files/figure-latex/cross-validation-clusters-2} }

@@

\hypertarget{cross-validating-model-selection}{%
\section{Cross-validating model
selection}\label{cross-validating-model-selection}}

As \citet[Sec. 7.10.2: ``The Wrong and Right Way to Do
Cross-validation'']{HastieTibshiraniFriedman:2009} explain, if the whole
data are used to select or fine-tune a statistical model, then
subsequent cross-validation of the model is intrinsically misleading,
because the model is selected to fit the whole data, including the part
of the data that remains when each fold is removed.

\hypertarget{a-preliminary-example}{%
\subsection{A preliminary example}\label{a-preliminary-example}}

The following example is similar in spirit to one employed by
\citet{HastieTibshiraniFriedman:2009}. Suppose that we randomly generate
\(n = 1000\) independent observations for a response variable variable
\(y \sim N(\mu = 10, \sigma^2 = 1)\), and independently sample \(1000\)
observations for \(p = 100\) ``predictors,'' \(x_1, \ldots, x_{100}\),
each from \(x_j \sim N(0, 1)\). The response has nothing to do with the
predictors and so the population linear-regression model
\(y_i = \alpha + \beta_1 x_{i1} + \cdots + \beta_{100} x_{i,100} + \varepsilon_i\)
has \(\alpha = 10\), all \(\beta_j = 0\), and
\(\sigma_\varepsilon^2 = 1\).

\begin{CodeChunk}
\begin{CodeInput}
R> set.seed(24361) # for reproducibility
R> D <- data.frame(
+   y = rnorm(1000, mean=10),
+   X = matrix(rnorm(1000*100), 1000, 100)
+ )
\end{CodeInput}
\end{CodeChunk}

Least-squares provides accurate estimates of the regression constant
\(\alpha = 10\) and the error variance \(\sigma^2 = 1\) for the ``null
model'' including only the regression constant; moreover, the omnibus
\(F\)-test of the correct null hypothesis that all of the \(\beta\)s are
0 for the ``full model'' with all 100 \(x\)s is associated with a large
\(p\)-value:

\begin{CodeChunk}
\begin{CodeInput}
R> m.full <- lm(y ~ ., data=D)
R> m.null <- lm(y ~ 1, data=D)
R> summary(m.null)
\end{CodeInput}
\begin{CodeOutput}

Call:
lm(formula = y ~ 1, data = D)

Residuals:
    Min      1Q  Median      3Q     Max 
-3.4585 -0.6809  0.0190  0.6365  2.9346

Coefficients:
            Estimate Std. Error t value Pr(>|t|)
(Intercept)  9.93704    0.03122   318.3   <2e-16

Residual standard error: 0.9872 on 999 degrees of freedom
\end{CodeOutput}
\begin{CodeInput}
R> anova(m.null, m.full)
\end{CodeInput}
\begin{CodeOutput}
Analysis of Variance Table

Model 1: y ~ 1
Model 2: y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7 + X.8 + X.9 + X.10 + 
    X.11 + X.12 + X.13 + X.14 + X.15 + X.16 + X.17 + X.18 + X.19 + 
    X.20 + X.21 + X.22 + X.23 + X.24 + X.25 + X.26 + X.27 + X.28 + 
    X.29 + X.30 + X.31 + X.32 + X.33 + X.34 + X.35 + X.36 + X.37 + 
    X.38 + X.39 + X.40 + X.41 + X.42 + X.43 + X.44 + X.45 + X.46 + 
    X.47 + X.48 + X.49 + X.50 + X.51 + X.52 + X.53 + X.54 + X.55 + 
    X.56 + X.57 + X.58 + X.59 + X.60 + X.61 + X.62 + X.63 + X.64 + 
    X.65 + X.66 + X.67 + X.68 + X.69 + X.70 + X.71 + X.72 + X.73 + 
    X.74 + X.75 + X.76 + X.77 + X.78 + X.79 + X.80 + X.81 + X.82 + 
    X.83 + X.84 + X.85 + X.86 + X.87 + X.88 + X.89 + X.90 + X.91 + 
    X.92 + X.93 + X.94 + X.95 + X.96 + X.97 + X.98 + X.99 + X.100
  Res.Df    RSS  Df Sum of Sq      F Pr(>F)
1    999 973.65                            
2    899 888.44 100    85.208 0.8622  0.825
\end{CodeOutput}
\end{CodeChunk}

Next, using the \code{stepAIC()} function in the \pkg{MASS} package
\citep{VenablesRipley:2002}, let us perform a forward stepwise
regression to select a ``best'' model, starting with the null model, and
using AIC as the model-selection criterion (see \code{?stepAIC()} for
details):\footnote{It's generally advantageous to start with the largest
  model, here the one with 100 predictors, and proceed by backward
  elimination. In this demonstration, however, where all of the
  \(\beta\)s are really 0, the selected model will be small, and so we
  proceed by forward selection from the null model to save computing
  time.}

\begin{CodeChunk}
\begin{CodeInput}
R> library("MASS")  # for stepAIC()
R> m.select <- stepAIC(m.null,
+                     direction="forward", trace=FALSE,
+                     scope=list(lower=~1, upper=formula(m.full)))
R> summary(m.select)
\end{CodeInput}
\begin{CodeOutput}

Call:
lm(formula = y ~ X.99 + X.90 + X.87 + X.40 + X.65 + X.91 + X.53 + 
    X.45 + X.31 + X.56 + X.61 + X.60 + X.46 + X.35 + X.92, data = D)

Residuals:
    Min      1Q  Median      3Q     Max 
-3.2620 -0.6446  0.0236  0.6406  3.1180

Coefficients:
            Estimate Std. Error t value Pr(>|t|)
(Intercept)  9.93716    0.03098 320.803  < 2e-16
X.99        -0.09103    0.03082  -2.953  0.00322
X.90        -0.08205    0.03135  -2.617  0.00901
X.87        -0.06942    0.03105  -2.235  0.02561
X.40        -0.04759    0.03076  -1.547  0.12211
X.65        -0.05523    0.03147  -1.755  0.07952
X.91         0.05245    0.03084   1.700  0.08937
X.53        -0.04921    0.03048  -1.615  0.10672
X.45         0.05543    0.03182   1.742  0.08183
X.31         0.04525    0.03108   1.456  0.14570
X.56         0.05433    0.03273   1.660  0.09723
X.61        -0.05085    0.03170  -1.604  0.10908
X.60        -0.05133    0.03194  -1.607  0.10832
X.46         0.05158    0.03272   1.576  0.11529
X.35         0.04696    0.03146   1.493  0.13584
X.92         0.04430    0.03100   1.429  0.15329

Residual standard error: 0.9725 on 984 degrees of freedom
Multiple R-squared:  0.04419,   Adjusted R-squared:  0.02962 
F-statistic: 3.033 on 15 and 984 DF,  p-value: 8.338e-05
\end{CodeOutput}
\begin{CodeInput}
R> mse(D$y, fitted(m.select))
\end{CodeInput}
\begin{CodeOutput}
[1] 0.9306254
attr(,"casewise loss")
[1] "(y - yhat)^2"
\end{CodeOutput}
\end{CodeChunk}

The resulting model has 15 predictors, a very modest \(R^2 = .044\), but
a small \(p\)-value for its omnibus \(F\)-test (which, of course, is
entirely spurious because the same data were used to select and test the
model). The MSE for the selected model is smaller than the true error
variance \(\sigma^2 = 1\), as is the estimated error variance for the
selected model, \(\widehat{\sigma}^2 = 0.973^2 = 0.947\).\footnote{The
  \code{"casewise loss"} attribute attached to the value returned by the
  \code{mse()} function signals to \code{cv()} that MSE is an average of
  casewise components (and how to calculate the casewise components),
  which has implications for computing bias correction and confidences
  intervals; see Section
  \ref{computation-of-the-bias-corrected-cv-criterion-and-confidence-intervals}.}

If we cross-validate the selected model, we also obtain an optimistic
estimate of its predictive power (although the confidence interval for
the bias-adjusted MSE includes 1):

\begin{CodeChunk}
\begin{CodeInput}
R> cv(m.select, seed=2529)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 2529
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 0.9593695
bias-adjusted cross-validation criterion = 0.9578478
95% CI for bias-adjusted CV criterion = (0.8766138, 1.039082)
full-sample criterion = 0.9306254 
\end{CodeOutput}
\end{CodeChunk}

The \code{cvSelect()} function in the \pkg{cv} package, in contrast,
allows us to cross-validate the whole model-selection procedure. The
first argument to \code{cvSelect()} is a model-selection function
capable of refitting the model with a fold omitted and returning a CV
criterion. The \code{selectStepAIC()} function, also in \pkg{cv} and
based on \code{stepAIC()}, is suitable for use with \code{cvSelect()}:

\begin{CodeChunk}
\begin{CodeInput}
R> cv.select <- cvSelect(selectStepAIC, data=D, seed=3791,
+                       model=m.null, direction="forward",
+                       scope=list(lower=~1, 
+                                  upper=formula(m.full)))
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 3791
\end{CodeOutput}
\begin{CodeInput}
R> cv.select
\end{CodeInput}
\begin{CodeOutput}
10-Fold Cross Validation
cross-validation criterion = 1.06873
bias-adjusted cross-validation criterion = 1.061183
95% CI for bias-adjusted CV criterion = (0.9717229, 1.150642)
full-sample criterion = 0.9306254 
\end{CodeOutput}
\end{CodeChunk}

The other arguments to \code{cvSelect()} are:

\begin{itemize}
\tightlist
\item
  \code{data}, the data set to which the model is fit.
\item
  \code{seed}, an optional seed for \proglang{R}'s pseudo-random-number
  generator; as for \code{cv()}, if the seed isn't supplied by the user,
  a seed is randomly selected and saved.
\item
  additional arguments required by the model-selection function, here
  the starting \code{model} argument, the \code{direction} of model
  selection, and the \code{scope} of models considered (from the model
  with only a regression constant to the model with all 100 predictors).
\end{itemize}

By default, \code{cvSelect()} performs 10-fold CV, and it produces an
estimate of MSE for the model-selection procedure even \emph{larger}
than the true error variance, \(\sigma^2 = 1\).

Also by default, when the number of folds is 10 or fewer,
\code{cvSelect()} saves the coefficients of the selected models. In this
example, the \code{compareFolds()} function reveals that the variables
retained by the model-selection process in the several folds are quite
different:
R> compareFolds(cv.select)
        (Intercept)    X.87    X.90    X.99    X.91    X.54    X.53    X.56
Fold 1       9.9187 -0.0615 -0.0994 -0.0942  0.0512  0.0516                
Fold 2       9.9451 -0.0745 -0.0899 -0.0614          0.0587          0.0673
Fold 3       9.9423 -0.0783 -0.0718 -0.0987  0.0601                  0.0512
Fold 4       9.9410 -0.0860 -0.0831 -0.0867  0.0570         -0.0508        
Fold 5       9.9421 -0.0659 -0.0849 -0.1004  0.0701  0.0511 -0.0487  0.0537
Fold 6       9.9633 -0.0733 -0.0874 -0.0960  0.0555  0.0629 -0.0478        
Fold 7       9.9279 -0.0618 -0.0960 -0.0838  0.0533         -0.0464        
Fold 8       9.9453 -0.0610 -0.0811 -0.0818          0.0497 -0.0612  0.0560
Fold 9       9.9173 -0.0663 -0.0894 -0.1100  0.0504  0.0524          0.0747
Fold 10      9.9449 -0.0745 -0.0906 -0.0891  0.0535  0.0482 -0.0583  0.0642
           X.40    X.45    X.65    X.68    X.92    X.15    X.26    X.46    X.60
Fold 1                  -0.0590                 -0.0456  0.0658  0.0608        
Fold 2                                   0.0607          0.0487                
Fold 3  -0.0496         -0.0664          0.0494                                
Fold 4  -0.0597  0.0579 -0.0531          0.0519 -0.0566                 -0.0519
Fold 5                           0.0587                          0.0527 -0.0603
Fold 6  -0.0596  0.0552          0.0474                                        
Fold 7           0.0572          0.0595                                        
Fold 8           0.0547 -0.0617  0.0453  0.0493 -0.0613  0.0591  0.0703 -0.0588
Fold 9  -0.0552  0.0573 -0.0635  0.0492         -0.0513  0.0484         -0.0507
Fold 10 -0.0558                          0.0529                  0.0710        
           X.61     X.8    X.28    X.29    X.31    X.35    X.70    X.89    X.17
Fold 1  -0.0490          0.0616 -0.0537                  0.0638                
Fold 2           0.0671                  0.0568                  0.0523        
Fold 3  -0.0631          0.0616                                                
Fold 4           0.0659         -0.0549          0.0527                  0.0527
Fold 5           0.0425                  0.0672  0.0613          0.0493        
Fold 6           0.0559         -0.0629  0.0498          0.0487                
Fold 7                                                           0.0611  0.0472
Fold 8  -0.0719                                          0.0586                
Fold 9                   0.0525                                                
Fold 10 -0.0580                                  0.0603                        
           X.25     X.4    X.64    X.81    X.97    X.11     X.2    X.33    X.47
Fold 1                                   0.0604          0.0575                
Fold 2   0.0478          0.0532  0.0518                                        
Fold 3                           0.0574                          0.0473        
Fold 4                   0.0628                                                
Fold 5   0.0518                                                                
Fold 6                                           0.0521                        
Fold 7           0.0550                                                        
Fold 9                                   0.0556                          0.0447
Fold 10          0.0516                                                        
            X.6    X.72    X.73    X.77    X.79 X.88
Fold 1   0.0476                                     
Fold 2                   0.0514                     
Fold 4                                  -0.0473     
Fold 5           0.0586                         0.07
Fold 6                          -0.0489             

\hypertarget{cross-validating-choice-of-transformations-in-regression}{%
\subsection{Cross-validating choice of transformations in
regression}\label{cross-validating-choice-of-transformations-in-regression}}

The \pkg{cv} package also provides a \code{cvSelect()} procedure,
\code{selectTrans()}, for choosing transformations of the predictors and
the response in regression.

Some background: As \citet[Sec. 8.2]{Weisberg:2014} explains, there are
technical advantages to having (numeric) predictors in linear regression
analysis that are themselves linearly related. If the predictors
\emph{aren't} linearly related, then the relationships between them can
often be straightened by power transformations. Transformations can be
selected after graphical examination of the data, or by analytic
methods, such as transforming the predictors towards multivariate
normality, which implies linearity. Once the relationships between the
predictors are linearized, it can be advantageous similarly to transform
the response variable towards normality.

Selecting transformations analytically raises the possibility of
automating the process, as would be required for cross-validation. One
could, in principle, apply graphical methods to select transformations
for each fold, but because a data analyst couldn't forget the choices
made for previous folds, the process wouldn't really be applied
independently to the folds.

To illustrate, we adapt an example appearing in several places in
\citet{FoxWeisberg:2019} (for example in Chapter 3 on transforming
data), using data on the prestige and other characteristics of 102
Canadian occupations circa 1970. The data are in the \code{Prestige}
data frame in the \pkg{carData} package:

\begin{CodeChunk}
\begin{CodeInput}
R> data("Prestige", package="carData")
R> summary(Prestige)
\end{CodeInput}
\begin{CodeOutput}
   education          income          women           prestige    
 Min.   : 6.380   Min.   :  611   Min.   : 0.000   Min.   :14.80  
 1st Qu.: 8.445   1st Qu.: 4106   1st Qu.: 3.592   1st Qu.:35.23  
 Median :10.540   Median : 5930   Median :13.600   Median :43.60  
 Mean   :10.738   Mean   : 6798   Mean   :28.979   Mean   :46.83  
 3rd Qu.:12.648   3rd Qu.: 8187   3rd Qu.:52.203   3rd Qu.:59.27  
 Max.   :15.970   Max.   :25879   Max.   :97.510   Max.   :87.20  
     census       type   
 Min.   :1113   bc  :44  
 1st Qu.:3120   prof:31  
 Median :5135   wc  :23  
 Mean   :5402   NA's: 4  
 3rd Qu.:8312            
 Max.   :9517            
\end{CodeOutput}
\end{CodeChunk}

The variables in the \code{Prestige} data set are:

\begin{itemize}
\tightlist
\item
  \code{education}: average years of education for incumbents in the
  occupation, from the 1971 Canadian Census.
\item
  \code{income}: average dollars of annual income for the occupation,
  from the Census.
\item
  \code{women}: percentage of occupational incumbents who were women,
  also from the Census.
\item
  \code{prestige}: the average prestige rating of the occupation on a
  0--100 ``thermometer'' scale, from a Canadian social survey conducted
  around the same time.
\item
  \code{type}, type of occupation, and \code{census}, the Census
  occupational code, which are not used in our example.
\end{itemize}

The object of a regression analysis for the \code{Prestige} data (and
their original purpose) is to predict occupational prestige from the
other variables in the data set.

The scatterplot matrix in Figure \ref{fig:scatterplot-matrix} (a)
(produced by the \code{scatterplotMatrix()} function in the \pkg{car}
package) for the numeric variables in the data reveals that the
distributions of \code{income} and \code{women} are positively skewed,
and that some of the relationships among the three predictors, and
between the predictors and the response (i.e., \code{prestige}), are
nonlinear:

{\centering \subfloat[\label{fig:scatterplot-matrix-1}]{\includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/scatterplot-matrix-1} }\subfloat[\label{fig:scatterplot-matrix-2}]{\includegraphics[width=0.5\linewidth]{JSS-article-reduced_files/figure-latex/scatterplot-matrix-2} }

\caption[Scatterplot matrix for the Prestige data, (a) untransformed and (b) with the predictors transformed towards multivariate normality]{Scatterplot matrix for the Prestige data, (a) untransformed and (b) with the predictors transformed towards multivariate normality.}\label{fig:scatterplot-matrix}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> library("car", quietly=TRUE)
R> scatterplotMatrix(~ prestige + income + education + women,
+                   data=Prestige, smooth=list(spread=FALSE))
\end{CodeInput}
\end{CodeChunk}

The \code{powerTransform()} function in the \pkg{car} package transforms
variables towards multivariate normality by a generalization of Box and
Cox's maximum-likelihood-like approach \citep{BoxCox:1964}. Several
``families'' of power transformations can be used, including the
original Box-Cox family, simple powers (and roots), and two adaptations
of the Box-Cox family to data that may include negative values and
zeros: the Box-Cox-with-negatives family and the Yeo-Johnson family; see
\citet[Chap. 8]{Weisberg:2014} and \citet[Chap. 3]{FoxWeisberg:2019} for
details. Because \code{women} has some 0 values, we use the Yeo-Johnson
family:

\begin{CodeChunk}
\begin{CodeInput}
R> trans <- powerTransform( cbind(income, education, women) ~ 1,
+                          data=Prestige, family="yjPower")
R> summary(trans)
\end{CodeInput}
\begin{CodeOutput}
yjPower Transformations to Multinormality 
          Est Power Rounded Pwr Wald Lwr Bnd Wald Upr Bnd
income       0.2678        0.33       0.1051       0.4304
education    0.5162        1.00      -0.2822       1.3145
women        0.1630        0.16       0.0112       0.3149

 Likelihood ratio test that all transformation parameters are equal to 0
                               LRT df      pval
LR test, lambda = (0 0 0) 15.73879  3 0.0012827
\end{CodeOutput}
\end{CodeChunk}

We thus have evidence of the desirability of transforming \code{income}
(by the \(1/3\) power) and \code{women} (by the \(0.16\) power---which
is close to the ``0'' power, i.e., the log transformation), but not
\code{education}. Applying the ``rounded'' power transformations makes
the predictors better-behaved (cf., Figures \ref{fig:scatterplot-matrix}
(a) and (b)):

\begin{CodeChunk}
\begin{CodeInput}
R> P <- Prestige[, c("prestige", "income", "education", "women")]
R> (lambdas <- trans$roundlam)
R> names(lambdas) <- c("income", "education", "women")
R> for (var in c("income", "education", "women")){
+   P[, var] <- yjPower(P[, var], lambda=lambdas[var])
+ }
R> scatterplotMatrix(~ prestige + income + education + women,
+                   data=P, smooth=list(spread=FALSE))
\end{CodeInput}
\end{CodeChunk}

Comparing the MSE for the regressions with the original and transformed
predictors shows an advantage to the latter:

\begin{CodeChunk}
\begin{CodeInput}
R> m.pres <- lm(prestige ~ income + education + women, data=Prestige)
R> m.pres.trans <- lm(prestige ~ income + education + women, data=P)
R> mse(Prestige$prestige, fitted(m.pres))
\end{CodeInput}
\begin{CodeOutput}
[1] 59.15265
attr(,"casewise loss")
[1] "(y - yhat)^2"
\end{CodeOutput}
\begin{CodeInput}
R> mse(P$prestige, fitted(m.pres.trans))
\end{CodeInput}
\begin{CodeOutput}
[1] 50.60016
attr(,"casewise loss")
[1] "(y - yhat)^2"
\end{CodeOutput}
\end{CodeChunk}

Similarly, component+residual plots for the two regressions in Figure
\ref{fig:CR-plots}, produced by the \code{crPlots()} function in the
\pkg{car} package, suggest that the partial relationship of
\code{prestige} to \code{income} is more nearly linear in the
transformed data, but the transformation of \code{women} fails to
capture what appears to be a slight quadratic partial relationship; the
partial relationship of \code{prestige} to \code{education} is close to
linear in both regressions:

\begin{CodeChunk}
\begin{CodeInput}
R> crPlots(m.pres)
R> crPlots(m.pres.trans)
\end{CodeInput}
\begin{figure}

{\centering \subfloat[\label{fig:CR-plots-1}]{\includegraphics[width=0.6\linewidth]{JSS-article-reduced_files/figure-latex/CR-plots-1} }\newline\subfloat[\label{fig:CR-plots-2}]{\includegraphics[width=0.6\linewidth]{JSS-article-reduced_files/figure-latex/CR-plots-2} }\newline

\caption[Component+residual plots for the Prestige regression with (a) the original predictors and (b) transformed predictors]{Component+residual plots for the Prestige regression with (a) the original predictors and (b) transformed predictors}\label{fig:CR-plots}
\end{figure}
\end{CodeChunk}

Having transformed the predictors towards multivariate normality, we now
consider whether there's evidence for transforming the response (using
\code{powerTransform()} for Box and Cox's original method), and we
discover that there's not:

\begin{CodeChunk}
\begin{CodeInput}
R> summary(powerTransform(m.pres.trans))
\end{CodeInput}
\begin{CodeOutput}
bcPower Transformation to Normality 
   Est Power Rounded Pwr Wald Lwr Bnd Wald Upr Bnd
Y1    1.0194           1       0.6773       1.3615

Likelihood ratio test that transformation parameter is equal to 0
 (log transformation)
                          LRT df       pval
LR test, lambda = (0) 32.2174  1 1.3785e-08

Likelihood ratio test that no transformation is needed
                             LRT df    pval
LR test, lambda = (1) 0.01238421  1 0.91139
\end{CodeOutput}
\end{CodeChunk}

The \code{selectTrans()} function in the \pkg{cv} package automates the
process of selecting predictor and response transformations. The
function takes a \code{data} set and ``working'' \code{model} as
arguments, along with the candidate \code{predictors} and
\code{response} for transformation, and the transformation \code{family}
to employ. If the \code{predictors} argument is missing then only the
\code{response} is transformed, and if the \code{response} argument is
missing, only the supplied \code{predictors} are transformed. The
default \code{family} for transforming the predictors is
\code{"bcPower"}---the original Box-Cox family---as is the default
\code{family.y} for transforming the response; here we specify
\code{family="yjPower"} because of the 0s in \code{women}.
\code{selectTrans()} returns the result of applying a lack-of-fit
criterion to the model after the selected transformation is applied,
with the default \code{criterion=mse}:

\begin{CodeChunk}
\begin{CodeInput}
R> selectTrans(data=Prestige, model=m.pres,
+             predictors=c("income", "education", "women"),
+             response="prestige", family="yjPower")
\end{CodeInput}
\begin{CodeOutput}
[1] 50.60016
attr(,"casewise loss")
[1] "(y - yhat)^2"
\end{CodeOutput}
\end{CodeChunk}

\code{selectTrans()} also takes an optional \code{indices} argument,
making it suitable for doing computations on a subset of the data (i.e.,
a CV fold), and hence for use with \code{cvSelect()} (see
\code{?selectTrans} and the vignette on extending the \pkg{cv} package
for details):

>>>>

{\centering \includegraphics[width=0.5\linewidth]{JSS-article-reduced-2_files/figure-latex/mpg-horsepower-scatterplot-polynomials-1}

@@

{\centering \includegraphics[width=0.5\linewidth]{JSS-article-reduced-2_files/figure-latex/mpg-horsepower-MSE-se-1}

@@

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 392419
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.16505
bias-adjusted cross-validation criterion = 19.15554
95% CI for bias-adjusted CV criterion = (15.71577, 22.59531)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> m.auto.glm <- glm(mpg ~ poly(horsepower, 2), data=Auto)
R> boot::cv.glm(Auto, m.auto.glm)$delta
\end{CodeInput}
\begin{CodeOutput}
[1] 19.24821 19.24787
\end{CodeOutput}
\begin{CodeInput}
R> microbenchmark::microbenchmark(
+   hatvalues = cv(m.auto, k="loo"),
+   Woodbury = cv(m.auto, k="loo", method="Woodbury"),
+   naive = cv(m.auto, k="loo", method="naive"),
+   cv.glm = boot::cv.glm(Auto, m.auto.glm),
+   times=10
+ )
\end{CodeInput}
\begin{CodeOutput}
Warning in microbenchmark::microbenchmark(hatvalues = cv(m.auto, k = "loo"), :
less accurate nanosecond times to avoid potential integer overflows
\end{CodeOutput}
\begin{CodeOutput}
Unit: microseconds
      expr        min         lq       mean     median         uq        max
 hatvalues    981.745   1172.477   1163.211   1189.943   1203.063   1325.817
  Woodbury   9958.777  10178.496  10309.593  10232.370  10356.067  10998.127
     naive 214083.345 217623.654 247100.940 220301.795 277217.933 331491.314
    cv.glm 381699.053 383011.832 401940.298 387351.887 439080.439 442256.709
 neval cld
    10 a  
    10 a  
    10  b 
    10   c
\end{CodeOutput}
\end{CodeChunk}

@@

{\centering \subfloat[\label{fig:polynomial-regression-CV-graph-2-1}]{\includegraphics[width=0.5\linewidth]{JSS-article-reduced-2_files/figure-latex/polynomial-regression-CV-graph-2-1} }\subfloat[\label{fig:polynomial-regression-CV-graph-2-2}]{\includegraphics[width=0.5\linewidth]{JSS-article-reduced-2_files/figure-latex/polynomial-regression-CV-graph-2-2} }

@@

{\centering \subfloat[\label{fig:plot1-1}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced-2_files/figure-latex/plot1-1} }\subfloat[\label{fig:plot1-2}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced-2_files/figure-latex/plot1-2} }\newline\subfloat[\label{fig:plot1-3}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced-2_files/figure-latex/plot1-3} }\subfloat[\label{fig:plot1-4}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced-2_files/figure-latex/plot1-4} }\newline\subfloat[\label{fig:plot1-5}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced-2_files/figure-latex/plot1-5} }

@@

{\centering \subfloat[\label{fig:cross-validation-clusters-1}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced-2_files/figure-latex/cross-validation-clusters-1} }\subfloat[\label{fig:cross-validation-clusters-2}]{\includegraphics[width=0.45\linewidth]{JSS-article-reduced-2_files/figure-latex/cross-validation-clusters-2} }

@@

\hypertarget{cross-validating-model-specification}{%
\section{Cross-validating model
specification}\label{cross-validating-model-specification}}

As \citet[Sec. 7.10.2: ``The Wrong and Right Way to Do
Cross-validation'']{HastieTibshiraniFriedman:2009} explain, if the whole
data are used to specify or fine-tune a statistical model, then
subsequent cross-validation of the model is intrinsically misleading,
because the model is selected to fit the whole data, including the part
of the data that remains when each fold is removed. Statistical modeling
is partly a craft, and one could imagine applying that craft to
successive partial data sets, each with a fold removed. The resulting
procedure would be tedious, though possibly worth the effort, but it
would also be difficult to realize in practice: After all, we can hardly
erase our memory of statistical modeling choices between analyzing
partial data sets. Alternatively, if we're able to automate the process
of model specification, then we can more realistically apply CV
mechanically.

The \code{cvSelect()} function in the \pkg{cv} package cross-validates a
model-specification process in a general manner. Functions for three
such model-specification processes are included in the package:
\code{selectStepAIC()}, based on the \code{stepAIC()} function in the
\pkg{MASS} package \citep{VenablesRipley:2002}, performs stepwise
predictor selection for regression models; \code{selectTrans()}, based
on the \code{powerTransform()} function in the \pkg{car} package
\citep{FoxWeisberg:2019}, transforms predictors and the response in a
regression model towards normality; and
\code{selectTransAndStepAIC()}---the use of which we illustrate in the
current section---performs both of these procedures sequentially. In the
vignette on extending the \pkg{cv} package, we explain how to add
model-selection procedures.

To illustrate cross-validation of model specification, we return to the
\code{Auto} data set:\footnote{This example benefits from an email
  conversation with Bill Venables, who of course isn't responsible for
  the use to which we've put his insightful remarks.}
